
```{r}
library(psych)
library(dplyr)
library(car)
library(stargazer)
```

```{r}
sp_student_data <- read.csv("~/Desktop/Sleep Paralysis/SPStudentStudy/sp_student_data_cleaned_final")
```

```{r}
sp_student_data <- sp_student_data[-1, ]
```

# PSQI Descriptive Stats
```{r}
# PSQI 
PSQI_mean <- mean(sp_student_data$PSQI)
PSQI_median <- median(sp_student_data$PSQI)
PSQI_sd <- sd(sp_student_data$PSQI)
PSQI_list <- list(
  Mean = PSQI_mean,
  Median = PSQI_median,
  SD = PSQI_sd
)
print(PSQI_list)
```

# Calcualate time average
```{r}
#Calculate time average
sp_bedtime <- (sp_student_data$Q11.3)

minutes_past_midnight <- sapply(strsplit(sp_bedtime, ":"), function(x) {
  hours <- as.numeric(x[1])
  minutes <- as.numeric(x[2])
  return(hours * 60 + minutes)
})

average_minutes <- mean(minutes_past_midnight)
median_minutes <- median(minutes_past_midnight)

average_hours <- floor(average_minutes / 60)
average_remaining_minutes <- average_minutes %% 60
average_time <- sprintf("%02d:%02d", as.integer(average_hours), as.integer(average_remaining_minutes))

median_hours <- floor(median_minutes / 60)
median_remaining_minutes <- median_minutes %% 60
median_time <- sprintf("%02d:%02d", as.integer(median_hours), as.integer(median_remaining_minutes))

```

# Calcualting wake up time
```{r} 
#Calcualting wake up time
sp_wakeup <- (sp_student_data_cleaned_LowSP$Q11.3)

minutes_past_midnight <- sapply(strsplit(sp_wakeup, ":"), function(x) {
  hours <- as.numeric(x[1])
  minutes <- as.numeric(x[2])
  return(hours * 60 + minutes)
})

average_minutes <- mean(minutes_past_midnight)
median_minutes <- median(minutes_past_midnight)
sd_minutes <- sd(minutes_past_midnight)

average_hours <- floor(average_minutes / 60)
average_remaining_minutes <- average_minutes %% 60
average_time <- sprintf("%02d:%02d", as.integer(average_hours), as.integer(average_remaining_minutes))

median_hours <- floor(median_minutes / 60)
median_remaining_minutes <- median_minutes %% 60
median_time <- sprintf("%02d:%02d", as.integer(median_hours), as.integer(median_remaining_minutes))

sd_hours <- floor(sd_minutes / 60)
sd_remaining_minutes <- sd_minutes %% 60
sd_time <- sprintf("%02d:%02d", as.integer(sd_hours), as.integer(sd_remaining_minutes))

time_list <- list(
  Mean = average_time,
  Median = median_time,
  SD = sd_time
)

print(time_list)
```

# Pearson's Correlation
```{r}
 # CORRELATION ANALYSIS
# Pearson's Correlation
sp_freq_model <- lm(Black_coded ~ number_of_occupants, data = sp_student_data_cleaned)
summary(sp_freq_model)
plot(sp_freq_model)
hist(resid(sp_freq_model))
```

# Spearman's R
```{r}
# CORRELATION ANALYSIS WITH GRAPH 
# Spearman's R 
spearman_corr <- cor.test(sp_student_data_cleaned$nicotine_factor_FA, sp_student_data_cleaned$alc_factor_FA, method = "spearman")
print(spearman_corr)
```

# Factor Analysis
```{r}
# FACTOR ANALYSIS
sp_factor_df <- sp_student_data_cleaned %>%
  select(napping_coded, napping_length_coded, alc_use_coded, alc_per_week, alc_at_once, alc_nights_per_week, caffetine_use_coded, caffetine_per_week, caffetine_per_day, family_members_coded, family_number_coded, nicotine_use_coded, nicotine_per_week, nicotine_per_day
  ) 
fa_result <- fa(sp_factor_df, nfactors = 5, rotate = "varimax")
print(fa_result)
fa.diagram(fa_result)
# Calculate FA scores 
sp_factor_scores_FA <- predict(fa_result, sp_factor_df)
head(sp_factor_scores_FA)
```

# Creating Factor Scores
```{r}
# CREATING TESTING FACTOR FOR FA
# Alcohol Factor 
sp_student_data_cleaned$alc_factor_FA <- sp_factor_scores_FA[,1]
# Caffeine Factor
sp_student_data_cleaned$caffetine_factor_FA <- sp_factor_scores_FA[,2]
# Family Factor
sp_student_data_cleaned$family_factor_FA <- sp_factor_scores_FA[,3]
# Nicotine Factor
sp_student_data_cleaned$nicotine_factor_FA <- sp_factor_scores_FA[,4]
# Napping Factor
sp_student_data_cleaned$napping_factor_FA <- sp_factor_scores_FA[,5]
```

# PCA with Varimax Rotation
```{r}
# PCA with Varimax Rotation
pca_result <- principal(sp_factor_df, nfactors = 4, rotate = "varimax")
print(pca_result)
fa.diagram(pca_result)
# Calculate factor scores
sp_factor_scores <- predict(pca_result, sp_factor_df)
head(sp_factor_scores)
```

# Creating PCA Factor Scores
```{r}
# CREATING THE FACTORS FOR PCA
# Caffeine Factor 
sp_student_data_cleaned$alc_factor <- sp_factor_scores[,1]
# Alcohol Factor 
sp_student_data_cleaned$caffetine_factor <- sp_factor_scores[,2]
# Family Factor
sp_student_data_cleaned$family_factor <- sp_factor_scores[,3]
# Napping Factor
sp_student_data_cleaned$napping_factor <- sp_factor_scores[,4]
```

# Cronbach's A
```{r}
# CRONBACH'S A

#subset each factor and it's variables
factor_for_alpha_analysis <- sp_student_data_cleaned[c("napping_coded", "napping_length_coded")]
#calculate the cronbach's a for the factor
cronbach_alpha_result <- alpha(factor_for_alpha_analysis)
print(cronbach_alpha_result)
```

# Hierarchical Linear Regression
```{r}
# HIERARCHICAL LINEAR REGRESSION
# Building the models
m0 <- lm(sp_freq_coded ~ 1, data = sp_student_data_cleaned_HighSP)
m1 <- lm(sp_freq_coded ~ age_coded + education_coded + is_female, data = sp_student_data_cleaned_HighSP)
m2 <- lm(sp_freq_coded ~ age_coded + education_coded + is_female + PSQI, data = sp_student_data_cleaned_HighSP)
m3 <- lm(sp_freq_coded ~ age_coded + education_coded + is_female + PSQI + caffetine_factor_FA + alc_factor_FA + napping_factor_FA + family_factor_FA, data = sp_student_data_cleaned_HighSP)
# anova comparison
anova(m0)
anova(m0, m1, m2, m3)
HGM_anova_results <- anova(m1, m2, m3)
# summary stats
summary(m1)
summary(m2)
summary(m3)
# checking assumptions
plot(m1, which = 1)
plot(m1, which = 2)
plot(sp_student_data_cleaned_HighSP$sp_freq_coded, residuals(m1))
vif(m1)
plot(m2, which = 1)
plot(m2, which = 2)
plot(sp_student_data_cleaned_HighSP$sp_freq_coded, residuals(m2))
vif(m2)
plot(m3, which = 1)
plot(m3, which = 2)
plot(sp_student_data_cleaned_HighSP$sp_freq_coded, residuals(m3))
vif(m3)

stargazer(HGM_anova_results, type = "text")
```

# Creation of High and Low SP Groups
```{r}
# CREATION OF HIGH AND LOW SP GROUPS
sp_student_data_cleaned_HighSP <- sp_student_data_cleaned %>%
  filter(sp_freq_coded >= 3)
sp_student_data_cleaned_LowSP <- sp_student_data_cleaned %>%
  filter(sp_freq_coded <= 2)
sp_student_data_cleaned <- sp_student_data_cleaned %>%
  mutate(sp_level = case_when(
    sp_freq_coded >= 3 ~ 1,
    sp_freq_coded <= 2 ~ 0
  ))
```

# Continous Variables Difference between High and Low SP
```{r}
# CON VARIABLES FOR DIFFERENCE WITH HIGH SP AND LOW SP
sp_time_bed <- (sp_student_data_cleaned_HighSP$nicotine_factor)
sp_time_bed <- as.numeric(sp_time_bed)
mean_value <- mean(sp_time_bed)
median_value <- median(sp_time_bed)
sd_variable <- sd(sp_time_bed)
con_list_high <- list(
  Mean_high = mean_value,
  Median_high = median_value,
  SD_high = sd_variable
)
print(con_list_high)
sp_time_bed <- (sp_student_data_cleaned_LowSP$nicotine_factor)
sp_time_bed <- as.numeric(sp_time_bed)
mean_value <- mean(sp_time_bed)
median_value <- median(sp_time_bed)
sd_variable <- sd(sp_time_bed)
con_list_low <- list(
  Mean_low = mean_value,
  Median_low = median_value,
  SD_low = sd_variable
)
print(con_list_low)
```

# Descriptive Stats for High and Low SP
```{r}
# DESCRIBIVE STATS FOR HIGH AND LOW SP FREQUENCY
sp_student_freq_table_high <- table(sp_student_data_cleaned_HighSP$tech_before_bed_coded) #just input the column
sp_student_percent_table_high <- round((prop.table(sp_student_freq_table_high)) * 100, digits = 2) 
sp_descriptive_table_high <- cbind(sp_student_freq_table_high, Percentage = sp_student_percent_table_high)
print(sp_descriptive_table_high)
sp_student_freq_table_low <- table(sp_student_data_cleaned_LowSP$tech_before_bed_coded) #just input the column
sp_student_percent_table_low <- round((prop.table(sp_student_freq_table_low)) * 100, digits = 2) 
sp_descriptive_table_low <- cbind(sp_student_freq_table_low, Percentage = sp_student_percent_table_low)
print(sp_descriptive_table_low)
```

# Mann-Whitney U Test
```{r}
# CHI SQUARED TEST, MANNU WHITNEY AND CRAMER'S V
sp_mannu_test <- wilcox.test(sp_student_data_cleaned_HighSP$education_coded, sp_student_data_cleaned_LowSP$education_coded, exact = FALSE)
print(sp_mannu_test)
combined_df_for_chi <- rbind(sp_student_data_cleaned_HighSP, sp_student_data_cleaned_LowSP)
table_data_for_chi <- table(combined_df_for_chi$sp_level, combined_df_for_chi$education_coded)
sp_chi_test <- chisq.test(table_data_for_chi)
n <- sum(table_data_for_chi)
k <- min(nrow(table_data_for_chi) - 1, ncol(table_data_for_chi) - 1)
cramer_v <- sqrt(sp_chi_test$statistic / (n * k))
print(sp_chi_test$statistic)
print(n)
print(k)
print(cramer_v)
```

# Shapiro-Wilk Test for Normality
```{r}
# Assuming 'data_vector' is your vector of data
shapiro_test_result <- shapiro.test(sp_student_data_cleaned$napping_factor_FA)

# Print the results
print(shapiro_test_result)
```

# Differences in Correlational Analysis with High and Low SP
```{r}
# DIFFERENCES IN CORRELATION ANALYSIS BETWEEN HIGH AND LOW SP
# Spearman's R
spearman_corr <- cor.test(sp_student_data_cleaned_HighSP$napping_factor, sp_student_data_cleaned_HighSP$sp_freq_coded, method = "spearman")
print(spearman_corr)

# Create a scatter plot with a regression line
ggplot(sp_student_data_cleaned_HighSP, aes(x = sp_freq_coded, y = napping_factor)) +
    geom_point() +
    geom_smooth(method = "lm", se = TRUE) +
    labs(title = "Correlation between Variable1 and Variable2",
         x = "Indepedent",
         y = "Dependent")

spearman_corr <- cor.test(sp_student_data_cleaned_LowSP$napping_factor, sp_student_data_cleaned_LowSP$sp_freq_coded, method = "spearman")
print(spearman_corr)

# Create a scatter plot with a regression line
ggplot(sp_student_data_cleaned_LowSP, aes(x = sp_freq_coded, y = napping_factor)) +
    geom_point() +
    geom_smooth(method = "lm", se = TRUE) +
    labs(title = "Correlation between Variable1 and Variable2",
         x = "Indepedent",
         y = "Dependent")

# OVERLAPPING GRAPHS WITH SP LEVEL AS A FACTOR
sp_student_data_cleaned$sp_level_factor <- factor(sp_student_data_cleaned$sp_level, 
                                           levels = c(0, 1), 
                                           labels = c("-SP", "+SP"))
p <- ggplot(sp_student_data_cleaned, aes(x = sp_freq_coded, y = napping_factor, color = sp_level_factor)) +
    geom_point() + 
    geom_smooth(method = "lm", aes(group = sp_level_factor), se = TRUE) +  
    scale_color_manual(values = c("-SP" = "blue", "+SP" = "red")) +
    labs(title = "Correlation of Sleep Paralysis Frequency and Outcome ",
         x = "Sleep Paralysis Frequency",
         y = "Outcome Variable") +
    theme_minimal() +
    theme(legend.position = "right")  # Adjust legend position

print(p)

line_plot <- ggplot(sp_student_data_cleaned) +
  stat_summary(
    mapping = aes(x = sp_level_factor, y = napping_factor), 
    fun.max = max,
    fun.min = min,
    fun = median
  )

print(line_plot)
```

# Fun Correlation Matrix for Variables
```{r}
# FUN CORRELATION MATRIX FOR THE VARIABLES
cor_matrix_fun <- cor(sp_student_data_cleaned[, c("napping_factor_FA", "number_of_occupants", "noisy_area_coded", "nicotine_factor", "alc_factor_FA", "caffetine_factor_FA", "family_factor_FA", "PSQI")], use = "complete.obs", method = "spearman")
print(cor_matrix_fun)
corrplot(cor_matrix_fun, method = "color", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, # Text label color and rotation
         diag = FALSE) # Exclude the diagonal
```

# Real Correlation Matrix for Variables
```{r}
# REAL CORRELATION MATRIx OF THE VARIABLES
# Compute the Spearman correlation matrix with p-values
# Compute the Spearman correlation matrix with p-values
rcorr_results <- rcorr(as.matrix(sp_student_data_cleaned[, c("napping_factor_FA", "number_of_occupants", "noisy_area_coded", "nicotine_factor", "alc_factor_FA", "caffetine_factor_FA", "family_factor_FA", "PSQI")]), type="spearman")

# Extract the matrix of correlations
cor_matrix <- rcorr_results$r

# Extract the matrix of p-values
p_matrix <- rcorr_results$P

# Format the correlation matrix with significance levels
formatted_cor_matrix <- mapply(function(x, p_value) {
  signif_level <- ifelse(p_value < 0.001, "***", 
                         ifelse(p_value < 0.01, "**", 
                                ifelse(p_value < 0.05, "*", "")))
  paste0(sprintf("%.2f", x), signif_level)
}, cor_matrix, p_matrix, SIMPLIFY = FALSE)

# Convert the list to a matrix and set the upper triangle to NA
formatted_cor_matrix <- matrix(formatted_cor_matrix, nrow = nrow(cor_matrix), byrow = TRUE,
                               dimnames = dimnames(cor_matrix))
formatted_cor_matrix[upper.tri(formatted_cor_matrix)] <- ""

# Display the formatted matrix using knitr::kable
knitr::kable(formatted_cor_matrix, format = "markdown", caption = "Spearman's rank correlation matrix with significance levels")

```

